---
title: "Data Science Homework 2"
output: github_document
---

The following libraries will be used throughout the homework. 

```{r load libraries, message = FALSE}
library(readxl)
library(tidyverse)
```


## Problem 1

Read and clean the Mr. Trash Wheel sheet according to homework specifications. 

```{r trash wheel dataset}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
```

Read and clean the precipitation data for 2017 and 2018 according to homework.

```{r precipitation datasets}
precipitation_18 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "2018 Precipitation", range = "A2:B15", col_names = TRUE) %>% 
  janitor::clean_names()

precipitation_17 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "2017 Precipitation", range = "A2:B15", col_names = TRUE) %>% 
  janitor::clean_names()
```

The Mr. Trash Wheel dataset consists of `r nrow(trash_wheel)` rows and `r ncol(trash_wheel)` columns. It includes the variables: `r names(trash_wheel)`. The median number of sports balls in a dumpster in 2017 is XX. 

## Problem 2

Read and clean the NYC transit data according to homework specifications. 

```{r NYC transit dataset, message = FALSE}
nyc_transit = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names()
```











