---
title: "Data Science Homework 2"
author: Kristina Howell
output: github_document
---

The following libraries will be used throughout the homework. 

```{r load libraries, message = FALSE}
library(readxl)
library(tidyverse)
```


## Problem 1

Read and clean the Mr. Trash Wheel sheet according to homework specifications. 

```{r trash wheel dataset}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))

#Another range argument: range = cell_cols("A:N")
```

Read and clean the precipitation data for 2018 and 2017 according to homework.

```{r precipitation datasets}
prec_18 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "2018 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  rename(prec = total) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
  
  #alternate argument, skip = 1, to skip the first row of table
  
prec_17 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "2017 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  rename(prec = total) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Merge the two datasets

```{r}
precip_df = bind_rows(prec_18, prec_17)


#Argument to convert numerical month into words

#month_df = 
  #tibble( 
    #month_num = 1:12, 
    #month_name = month.name)

#left_join(precip_df, month_df, by = "month")
```


The Mr. Trash Wheel dataset consists of `r nrow(trash_wheel)` rows and `r ncol(trash_wheel)` columns. It includes the variables: `r names(trash_wheel)`. The median number of sports balls in a dumpster in 2017 is XX. 

## Problem 2

Read and clean the NYC transit data according to homework specifications. 

```{r NYC transit dataset, message = FALSE}
nyc_transit = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, entrance_type, vending, ada)
```

Convert the entry variable from character to logical

```{r}

```

The nyc_transit database contains the variables: `r names(nyc_transit)`. It contains `r nrow(nyc_transit)` rows and `r ncol(nyc_transit)` columns. 

The data cleaning steps have included cleaning the names, using the clean_names() function, and selecting the variables to keep in the data set, using gthe select() function. I then converted the character variable, entry, into a logical vector, using ...

The data is mostly tidy, but the columns regarding route number could be improved. 

There are `r nrow(distinct(nyc_transit, line, station_name))` distinct stations within the data set. Of these `r nrow(distinct(nyc_transit, line, station_name))` stations, `r nrow(distinct(nyc_transit, station_name, ada = "TRUE"))`


## Problem 3






