---
title: "Data Science Homework 2"
author: Kristina Howell
output: github_document
---

The following libraries will be used throughout the homework. 

```{r load_libraries, message = FALSE}
library(readxl)
library(tidyverse)
```


## Problem 1

Read and clean the Mr. Trash Wheel sheet according to homework specifications. 

```{r trash_wheel_dataset}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls), 
         sports_balls = as.integer(sports_balls))

#Another range argument: range = cell_cols("A:N")
```

Read and clean the precipitation data for 2018 and 2017 according to homework.

```{r precipitation_datasets}
prec_18 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "2018 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  rename(prec = total) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
  
  #alternate argument, skip = 1, to skip the first row of table
  
prec_17 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                         sheet = "2017 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  rename(prec = total) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Merge the two datasets

```{r binding_datasets}
precip_df = bind_rows(prec_18, prec_17)


#Argument to convert numerical month into words

#month_df = 
  #tibble( 
    #month_num = 1:12, 
    #month_name = month.name)

#left_join(precip_df, month_df, by = "month")
```


The Mr. Trash Wheel dataset consists of `r nrow(trash_wheel)` rows and `r ncol(trash_wheel)` columns. It includes the variables: `r names(trash_wheel)`. The median number of sports balls in a dumpster in 2017 is `r median(trash_wheel$sports_balls)` .

## Problem 2

Read and clean the NYC transit data according to homework specifications. 

```{r NYC_transit_dataset, message = FALSE}
nyc_transit = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, 
         entry, entrance_type, vending, ada) %>% 
  mutate(ada = as.logical(ada))
```

The nyc_transit dataset offers general but comprehensive information about the transit stations located in NYC. It contains the variables: `r names(nyc_transit)`. Overall, the dataset has `r nrow(nyc_transit)` rows and `r ncol(nyc_transit)` columns. 

The data cleaning steps have included cleaning the names, using the clean_names() function, selecting the variables to keep in the data set, using the select() function, and converting the ada character variable to a logical variable using the mutate() function. 

The data is mostly tidy, but the columns regarding route number could be improved. 

#### Problem 2 Questions

###### How many distinct stations are there? 

There are `r nrow(distinct(nyc_transit, line, station_name))` distinct stations within the data set. 

###### How many stations are ADA compliant?

To determine which distinct stations are ADA compliant, run the following code chunk. 

```{r ada}
ada = filter(nyc_transit, ada == "TRUE") %>% 
   distinct(line, station_name)
```

Results will output in a tibble that lists line and station name, with a total of `r nrow(distinct(ada, line, station_name))` observations (stations). 

###### What proportion of station entrances / exits without vending allow entrance?

To determine which distinct stations do not have vending but do allow entry, apply the following code chunk. 

```{r vend}
vend_entry = filter(nyc_transit, vending == "NO", entry == "YES") %>% 
   distinct(line, station_name)
```

Results will output in a tibble that lists line and station name, with a total of `r nrow(distinct(vend_entry, line, station_name))` observations (stations). 

##### Reformat data so that route number and route name are distinct variables. 

```{r}

```


###### How many distinct stations serve the A train? 

###### Of the stations that serve the A train, how many are ADA compliant?



## Problem 3

Import the data sets

```{r import_data, message = FALSE}

pols_month = read_csv("./Data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day"))
  
unemploy = read_csv("./Data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names()
  
snp = read_csv("./Data/fivethirtyeight_datasets/snp.csv")

```



